A Decision Procedure for (Co)datatypes in SMT Solvers

### 2014-12-14 tex/codatadp.tex
more eval

To evaluate the decision procedure, we generated problems from existing
Isabelle formalizations using Sledgehammer \cite{paulson-blanchette-2010}.
Codatatypes being a recent addition to Isabelle
\cite{blanchette-et-al-2014-impl}, benchmarks are somewhat hard to come by. We
included the theory files from the Isabelle distribution and the \emph{Archive
of Formal Proofs} (AFP) \cite{klein-et-al-afp} that define codatatypes falling
within the supported fragment. We also included two unpublished theories by
Andreas Lochbihler and Peter Gammie. We complemented the benchark suite with
some theories about various list and tree data structures, to also test the
datatypes. The theory files were selected before running the experiments.

  * for each goal, select 256 lemmas as axioms, before monomorphization
    (after monomorphization, typically more formulas due to duplicates),
    and encode with the goal

  * then ran CVC4 for XXX seconds on the StarExec cluster \cite{xxx}

  * wanted to evaluate both the impact of datatypes and codatatypes
  * datatypes are known to be very useful on SMT-LIB, e.g. blah
  * but we wanted to see how much they make a difference in Isabelle

  * ran all problems four times, to try all four combinations of datatypes and
    codatatypes enabled or disabled

@@

\begin{table}[t!]
\begin{center}\begin{tabular}{l@{\kern1.5em}c@{\kern.5em}c@{\kern.5em}c@{\kern1.5em}c}
  & \HD{Distro} & \HD{AFP} & \HD{\landg} & All
\\[1pt] %%% TYPESETTING HACK
\midrule
%$\enatT$
\\[-9pt] %%% TYPESETTING HACK
No decision procedure
  & 800 & 922 & 450 & 2172 \\
Codatatypes without bisimilarity
  & 810 & 936 & 470 & 2216 \\
Codatatypes with bisimilarity
  & 811 & 939 & 472 & 2222 \\
Datatypes without cyclicity
  & 815 & 947 & 469 & 2231 \\
Datatypes with cyclicity
  & 818 & 953 & 470 & 2241 \\
Full decision procedure
  & 840 & 970 & 490 & 2300 \\[\jot]
Total number of goals
  & 1211\phantom{0} & 1121\phantom{0} & 899 & 3001
\end{tabular}\end{center}
\caption{Number of solved goals for three benchmark suites}
\end{table}

>>>>

To evaluate the decision procedure, we generated problems from existing
Isabelle formalizations using Sledgehammer \cite{paulson-blanchette-2010}.
Codatatypes being a recent addition to Isabelle
\cite{blanchette-et-al-2014-impl}, benchmarks are somewhat hard to come by. We
included the theory files from the Isabelle distribution and the \emph{Archive
of Formal Proofs} (AFP) \cite{klein-et-al-afp} that define codatatypes falling
within the supported fragment. We also included two unpublished theories by
Andreas Lochbihler and Peter Gammie. We complemented the benchmarks with some
theories about various list and tree data structures, to also test the
datatypes. The theory files were selected before running the experiments. Our
experimental data is publicly available \cite{our-eval-data}.

For each goal in each theory file, Sledgehammer was invoked to select
256~lemmas as axioms, which were then monomorphized and translated to SMT-LIB
along with the goal \cite{boehme-2012-phd}. The resulting problem was given to
CVC4, running for up to \textbf{TODO: XXX} seconds on the StarExec cluster
\cite{xxx}.

  * ran all problems several times, to try all four combinations of datatypes
    and codatatypes enabled or disabled -- also, for datatypes, we tried with
    and without the cyclicity rule; and for codatatypes, with and without the
    bisimilarity rule, so find out whether these somewhat expensive rules are
    useful in practice

@@

\begin{table}[t!]
\begin{center}\begin{tabular}{l@{\kern1.5em}c@{\kern.5em}c@{\kern.5em}c@{\kern1.5em}c}
  & \HD{Distro} & \HD{AFP} & \HD{\landg} & All
\\[1pt] %%% TYPESETTING HACK
\midrule
%$\enatT$
\\[-9pt] %%% TYPESETTING HACK
No decision procedure
  & 800 & 922 & 450 & 2172 \\
Codatatypes without bisimilarity
  & 810 & 936 & 470 & 2216 \\
Codatatypes with bisimilarity
  & 811 & 939 & 472 & 2222 \\
Datatypes without cyclicity
  & 815 & 947 & 469 & 2231 \\
Datatypes with cyclicity
  & 818 & 953 & 470 & 2241 \\
Full decision procedure
  & 840 & 970 & 490 & 2300 \\[\jot]
Total number of goals
  & 1211\phantom{0} & 1121\phantom{0} & 899 & 3001
\end{tabular}\end{center}
\caption{Number of solved goals for the three benchmark suites}
\end{table}

* compare with BBP 2011!! --- potentially makes a more significant difference
  than arithmetic (although the benchmark suites are not directly comparable)
